---
title: "in_game_nhl_win_model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCurl); library(xml2); library(rvest); library(jsonlite); library(foreach)
library(lubridate)
library(tidyverse)
source("EH_scrape_functions.R")
```


## Modelling ingame nhl win probabilities

I want to track the probability of the home team winning while the game is happening. To do this I will use Logstic regressiont to model Pr(Home team winning).

## Brief primer on Logistic Regresion

Logistic regression is a statisical model that models binary variables, which makes it perfect for our problem since there are only 2 outcomes to 
an NHL game (I am considering an OT loss to be a loss).

## Downloading the Data
```{r message=FALSE}
events_summary <- read_csv("data/events_summary_df.csv")
game_info <- read_csv("data/game_info_df.csv")
pbp_base <- read_csv("data/pbp_base.csv")
pbp_extras <- read_csv("data/pbp_extras.csv")
player_periods <- read_csv("data/player_periods.csv")
player_shifts <- read_csv("data/player_shifts.csv")
report <- read_csv("data/report.csv")
roster <- read_csv("data/roster_df.csv")
scratches <- read_csv("data/scratches_df.csv")
```

The data was collected using Evolving Hockey's scraper (put link to Evolving Hockey's scraper github)


## Creating the model

Taking the raw data, we have to transform it to extract the variables of interest and make it ready to be put into our model. The variables we will be using are:

* seconds into the game
* goals for
* goals against
* number of away players on the ice
* number of home players on the ice


What sort of data are we going to need to?
  seconds into the game
  home strength state (num players home - num players away)
  
  -Not needed
  event type (dummy variables), should also indicate if it's home or away event type 
    goal
    blocked shot
    shot (xG might be a better indicator)
    giveaway
    takeaway
    faceoff
    
  home and away players on ice (dummy var)
  teams (dummy var)
  
  
```{r}
# view(new_pbp %>% filter(game_id == gid))

data_pipeline <- function(pbp){
  cols_to_select <- c("game_id", "game_seconds","home_team", "away_team",
                      "home_score", "away_score", "game_strength_state", "event_type")
  
  events_to_remove <- c("GSTART", "GEND", "PSTART", "PEND")
  
  new_pbp <- pbp %>% 
              filter(game_seconds > 0 & !event_type %in% events_to_remove) %>% 
              dplyr::select(all_of(cols_to_select)) %>%
              mutate(GF = home_score) %>%
              mutate(GA = away_score) %>%
              mutate(num_home_players = str_extract(game_strength_state, "^.")) %>%
              mutate(num_away_players = str_extract(game_strength_state, ".$")) %>%
              mutate(num_home_players = ifelse(num_home_players == "E", 
                                               ifelse(game_seconds < 3600, "5", "3"), 
                                               num_home_players)) %>%
              mutate(num_away_players = ifelse(num_away_players == "E", 
                                               ifelse(game_seconds < 3600, "5", "3"),
                                               num_away_players)) %>%
              mutate(num_home_players = as.numeric(num_home_players)) %>%
              mutate(num_away_players = as.numeric(num_away_players))
  # dmy1 <- dummyVars("~home_team", data=new_pbp, fullRank = F)
  # dmy2 <- dummyVars("~away_team", data=new_pbp, fullRank = F)
  # 
  # home_team_dummy <- as_tibble(predict(dmy1, new_pbp))
  # away_team_dummy <- as_tibble(predict(dmy2, new_pbp))
  # 
  # new_pbp <- bind_cols(new_pbp, home_team_dummy, away_team_dummy)
  
  new_pbp <- new_pbp %>% dplyr::select(-c("home_team", "away_team", 
                                          "event_type", "game_strength_state",
                                          "home_score", "away_score"))
  new_pbp <- type_convert(new_pbp) 
}

new_pbp <- data_pipeline(pbp_base)
new_pbp %>% filter(game_id == "2018020001")
```

## Logistic Regression Model
```{r}
y <- game_info %>% mutate(home_team_win=ifelse(home_score>away_score, 1, 0)) %>%
                  dplyr::select("game_id","home_team_win")
df <- inner_join(y, new_pbp)
model <- glm(home_team_win~., data=df %>% dplyr::select(-c("game_id")), family="binomial")
summary(model)

saveRDS(model, "ingame_win_prob_model.RDS")
```

## RT Odds graph for Logistic Regression
```{r}
rt_odds_graph <- function(model, pbp, game_info){
  new_pbp <- data_pipeline(pbp)
  odds_home_team_winning <- exp(predict(model, newdata = new_pbp %>% dplyr::select(-c("game_id"))))
  # prob_home_team_winging <- predict(model, newdata = new_pbp %>% dplyr::select(-c("game_id")), type="response")
  new_pbp <- cbind(new_pbp, odds_home_team_winning)
  print(new_pbp)
  ggplot(new_pbp, aes(x=game_seconds, y=odds_home_team_winning)) + 
    geom_line() + ggtitle(paste("In Game Odds Tracker ", game_info$away_team, " @ ", 
                                game_info$home_team, ", ", game_info$game_date, sep="")) +
    theme(plot.title = element_text(hjust = 0.5))

}

rt_probs_graph <- function(model, pbp, game_info){
  new_pbp <- data_pipeline(pbp)
  prob_home_team_winging <- predict(model, newdata = new_pbp %>% dplyr::select(-c("game_id")), type="response")
  new_pbp <- cbind(new_pbp, prob_home_team_winging)
  print(new_pbp)
  ggplot(new_pbp, aes(x=game_seconds, y=prob_home_team_winging)) + 
    geom_line() + ggtitle(paste("In Game Win Probability Tracker ", game_info$away_team, " @ ", 
                                game_info$home_team, ", ", game_info$game_date, sep="")) +
    theme(plot.title = element_text(hjust = 0.5))
}
gid <- "2018020005"
df2 <- sc.scrape_pbp(games = c("2019030054"))
# rt_odds_graph(model, pbp_base %>% dplyr::filter(game_id == gid), game_info %>% dplyr::filter(game_id == gid))
rt_odds_graph(model, df2$pbp_base, df2$game_info_df)
rt_probs_graph(model, df2$pbp_base, df2$game_info_df)
```



```{r}
sched <- sc.scrape_schedule(start_date = Sys.Date()-20, end_date=Sys.Date()) %>% filter(away_team == "FLA" | home_team == "FLA")
sched

df3 <- sc.scrape_pbp(games=c("2019030044"))
rt_probs_graph(model, df3$pbp_base, df3$game_info)
```





