---
title: "in_game_nhl_win_model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCurl); library(xml2); library(rvest); library(jsonlite); library(foreach)
library(lubridate)
library(tidyverse)
library(googledrive)
library(arm)
library(caret)
library(glmnet)
```


Modelling ingame nhl win probabilities

want to model P(home team win) while the game is happening

Downloading the Data
```{r message=FALSE}
events_summary <- read_csv("data/events_summary_df.csv")
game_info <- read_csv("data/game_info_df.csv")
pbp_base <- read_csv("data/pbp_base.csv")
pbp_extras <- read_csv("data/pbp_extras.csv")
player_periods <- read_csv("data/player_periods.csv")
player_shifts <- read_csv("data/player_shifts.csv")
report <- read_csv("data/report.csv")
roster <- read_csv("data/roster_df.csv")
scratches <- read_csv("data/scratches_df.csv")
```

Getting shift data

What sort of data are we going to need to?
  seconds into the game
  home strength state (num players home - num players away)
  event type (dummy variables), should also indicate if it's home or away event type
    goal
    blocked shot
    shot (xG might be a better indicator)
    giveaway
    takeaway
    faceoff
  home and away players on ice (dummy var)
  teams (dummy var)
  
  
```{r}

data_pipeline_2 <- function(pbp){
  event_types <- c("BLOCK", "SHOT", "GIVE", "FAC", "HIT")
  cols_to_select <- c("game_id", "game_seconds", "event_type",
                      "event_team", "home_team", "away_team",
                      "home_score", "away_score", "game_strength_state", 
                      "game_score_state")
}

data_pipeline <- function(pbp){
  event_types <- c("BLOCK", "SHOT", "GIVE", "FAC", "HIT")
  cols_to_select <- c("game_id", "game_seconds", "event_type",
                      "event_team", "home_team", "away_team",
                      "home_score", "away_score", "game_strength_state", 
                      "game_score_state")
  
  new_pbp <- pbp %>% 
              dplyr::filter(game_seconds > 0 & event_type %in% event_types) %>% 
              dplyr::select(all_of(cols_to_select)) %>%
              dplyr::mutate(GF = as.numeric(str_extract(game_score_state, "^\\d"))) %>%
              dplyr::mutate(GA = as.numeric(str_extract(game_score_state, "\\d$"))) %>%
              dplyr::mutate(num_home_players = as.numeric(str_extract(game_strength_state, "^\\d"))) %>%
              dplyr::mutate(num_away_players = as.numeric(str_extract(game_strength_state, "\\d$"))) %>%
              dplyr::mutate(GFA = GF - GA) %>%
              dplyr::mutate(homeaway_players = num_home_players - num_away_players) %>%
              dplyr::mutate(event_team = ifelse(event_team == home_team, 1, 0)) %>%
              drop_na()
  
  dummy_model <- dummyVars("~event_type", data=new_pbp)
  event_type_dummy <- data.frame(predict(dummy_model, newdata=new_pbp))
  new_pbp <- cbind(new_pbp, event_type_dummy) %>% as_tibble(.)
  
  new_pbp <- new_pbp %>% dplyr::select("game_id", "game_seconds", "event_team", 
                            "homeaway_players", "event_typeSHOT", "event_typeGIVE",
                            "event_typeBLOCK","GF", "GA", "event_typeHIT")
  
  new_pbp <- type_convert(new_pbp) 
}

new_pbp <- data_pipeline(pbp_base)
```

```{r}
y <- game_info %>% mutate(home_team_win=ifelse(home_score>away_score, 1, 0)) %>% dplyr::select("game_id", "home_team_win")
df <- inner_join(y, new_pbp)
model <- glm(home_team_win~., data=df %>% dplyr::select(-"game_id"), family="binomial")
summary(model) 
```


```{r}
rt_odds_graph <- function(model, pbp, game_info){
  new_pbp <- data_pipeline(pbp)
  
  prob_home_team_winning <- predict(model, newdata = new_pbp %>% dplyr::select(-c("game_id")), type="response")
  new_pbp <- cbind(new_pbp, prob_home_team_winning)
  print(new_pbp)
  ggplot(new_pbp, aes(x=game_seconds, y=prob_home_team_winning)) + 
    geom_line() + ggtitle(paste("In Game Win Probability Tracker ", game_info$away_team, " @ ", 
                                game_info$home_team, ", ", game_info$game_date, sep="")) +
    theme(plot.title = element_text(hjust = 0.5))

}
gid <- "2018020153"
rt_odds_graph(model, pbp_base %>% dplyr::filter(game_id == gid), game_info %>% dplyr::filter(game_id == gid))
```


