---
title: "A Mixed Effect Model approach to Expected Goals in Hockey using R-INLA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}
library(INLA)
library(ggridges)
library(tidyverse)
library(reshape2)
library(parallel)
numCores <- detectCores()
pbp_data <- read_rds("../data/pbp_base.RDS")
sub_pbp <- pbp_data %>% filter(season == 20202021)
```



# Introduction

In this article, I will be talking about how we can use a Hierarchical model to model expected goals in hockey. If you are wondering what is a multilevel model, or what is an expected goal, don't worry I will be going over what these mean. The goal for this article is to quantify how different factors affect the odds of a shot becoming a goal using

# What is an Expected Goal?

# Mixed Effect Model

Mixed Effect Model, hierarchical model, multilevel models, they go by many names but they all essentially do the same thing, try to model data that has some grouping or nesting structure. It's an extension of linear regression that contains both random and fixed effects. They can accounts for non independence in the data. In this case, we are looking at NHL shot data, so each observation of our data is a shot and various information about the shot like where the shot took place, the distance of the shot from the net, the player who took the shot, the goalie and more. One of the ways we can group our data is according to the player that took the shot, since our data consists of players who have taken multiple shots. 

I have elected to model this from a Bayesian perspective. I find mixed effect models make more sense from this perspective and can be more numerically stable.

Since our data is also binary, we will be using a logistic mixed model which can be defined as follows using lme4 notation:

`isgoal ~ x + y + shot_distance + shot_angle + seconds_since_last_shot + (1|goalie) + (1|shooter) + (1|event_detail) + (1|score_state) + (1|game_strength_state)`

Since this is a Bayesian model, we need priors. On the fixed effects, we put a $ \mathcal{N}(0, 1)$ prior and the random effects will each have a $\mathcal{N}(0, \sigma)$ distribution and the prior for the $\sigma$ will be _________.

# INLA

The most common way to perform Bayesian inference is to use a library to perform MCMC like Stan but since our we have a lot of data, doing MCMC would take a very long time. We can take an alternative and use a method called Integrated Nested Laplace Approximation (INLA). 

# Putting it all together

To start with, we take the play by play data and we filter all the rows that are shots and select the variables we are interested. We put all this in a function which will help when we apply our model to different seasons.

```{r}
data_pipeline_inla_xG <- function(pbp){
  new_pbp <- pbp %>% filter(event_type %in% c("SHOT", "GOAL", "MISS") &
                              game_strength_state %in% c("5v5", "5v4", "4v5",
                                                         "4v4", "3v3")) %>%
    mutate(home_score_state = home_score - away_score,
           away_score_state = away_score - home_score,
           tied = if_else(event_team == home_team,
                          1*(home_score_state == 0), 1*(away_score_state == 0)),
           up_by_1 = if_else(event_team == home_team,
                          1*(home_score_state == 1), 1*(away_score_state == 1)),
           up_by_2 = if_else(event_team == home_team,
                          1*(home_score_state == 2), 1*(away_score_state == 2)),
           up_by_more_than_2 = if_else(event_team == home_team,
                          1*(home_score_state > 2), 1*(away_score_state > 2)),
           down_by_1 = if_else(event_team == home_team,
                          1*(home_score_state == -1), 1*(away_score_state == -1)),
           down_by_2 = if_else(event_team == home_team,
                          1*(home_score_state == -2), 1*(away_score_state == -2)),
           down_by_more_than_2 = if_else(event_team == home_team,
                          1*(home_score_state < -2), 1*(away_score_state < -2)),
           goalie = ifelse(event_team == home_team, home_goalie, away_goalie),
           shooter = event_player_1,
           y = coords_y,
           x = abs(coords_x),
           shot_distance = sqrt((89-x)^2 + y^2),
           shot_angle = atan2(abs(y), abs(89-x)),
           is_home_team = if_else(event_team == home_team, 1, 0),
           seconds_since_last_shot = ifelse(game_seconds - lag(game_seconds) < 0,
                                            game_seconds, game_seconds - lag(game_seconds)),
           isgoal=ifelse(event_type == "GOAL", 1, 0)) %>%
    pivot_longer(cols=all_of(c("tied", "up_by_1", "up_by_2", "up_by_more_than_2",
                               "down_by_1", "down_by_2", "down_by_more_than_2")),
                 names_to = "score_state", values_to = "count") %>%
    filter(count == 1) %>%
    dplyr::select("x", "y", "isgoal", "game_id", "game_seconds",
                  "shooter", "goalie", "shot_distance",
                  "shot_angle", "seconds_since_last_shot",
                  "event_team", "is_home_team", "event_detail",
                  "game_strength_state", "score_state", "game_seconds") %>% 
    drop_na() %>%
    mutate(shooter = ifelse(Encoding(shooter) != "unknown", enc2native(shooter),
                          shooter),
           shooter_id = vctrs::vec_group_id(shooter),
           goalie_id = vctrs::vec_group_id(goalie),
           event_detail_id = vctrs::vec_group_id(event_detail),
           score_state_id = vctrs::vec_group_id(score_state))
    
      
  return(new_pbp)
}
xg_data <- data_pipeline_inla_xG(sub_pbp) 

xg_data$shooter %>% vctrs::vec_group_id()
```

```{r}
# INLA model
prior_random <- list(prec = list(prior = "normal", param = c(0, 1)))
prior_fixed <- list(mean.intercept = 0, prec.intercept = 1,
                    mean = list(x = 0, y = 0, shot_distance = 0,
                                shot_angle = 0, seconds_since_last_shot = 0),
                    prec = list(x = 1, y = 1, shot_distance = 1,
                                shot_angle = 1, seconds_since_last_shot = 1))

f <- isgoal ~ 1 + x + y + shot_distance + shot_angle + seconds_since_last_shot +
  f(goalie, model="iid", hyper = prior_random) +
  f(shooter, model="iid",  hyper = prior_random) +
  f(event_detail, model="iid",  hyper = prior_random) +
  f(score_state, model="iid", hyper = prior_random) +
  f(game_strength_state, model="iid", hyper = prior_random)

inla_xg <- inla(f, data=xg_data,
                    family="binomial",
                control.predictor = list(link = 1),
                control.fixed = prior_fixed,
                control.compute = list(config=TRUE))
summary(inla_xg)
```


```{r}
# Stan model
players <- xg_data %>% pull(shooter)

```




```{r}
posterior <- function(n_samples, inla_model){
  selection <- list("shooter"=1:(inla_model$summary.random$shooter %>% nrow()),
    "goalie"=1:(inla_model$summary.random$goalie %>% nrow()),
    "event_detail"=1:(inla_model$summary.random$event_detail %>% nrow()),
    "score_state"=1:(inla_model$summary.random$score_state %>% nrow()),
    "game_strength_state"=1:(inla_model$summary.random$game_strength_state %>%
                               nrow()),
    "x"=1, "y"=1, "shot_distance"=1, "shot_angle"=1,
    "seconds_since_last_shot"=1, "(Intercept)"=1)

  rownames <- bind_cols(
    old_vars = c(
      paste("shooter",
            1:(inla_model$summary.random$shooter %>% nrow()), sep=":"),
      paste("goalie",
            1:(inla_model$summary.random$goalie %>% nrow()), sep=":"),
      paste("event_detail",
            1:(inla_model$summary.random$event_detail %>% nrow()), sep=":"),
      paste("score_state",
            1:(inla_model$summary.random$score_state %>% nrow()), sep=":"),
      paste("game_strength_state",
            1:(inla_model$summary.random$game_strength_state %>% nrow()), sep=":"),
          "x:1", "y:1", "shot_distance:1", "shot_angle:1",
      "seconds_since_last_shot:1", "(Intercept):1"
    ),
    new_vars = c(
      inla_model$summary.random$shooter$ID,
      inla_model$summary.random$goalie$ID,
      inla_model$summary.random$event_detail$ID,
      inla_model$summary.random$score_state$ID,
      inla_model$summary.random$game_strength_state$ID,
       "x", "y", "shot_distance", "shot_angle",
      "seconds_since_last_shot", "(Intercept)"
    )
  )
  inla_xg_posterior <- inla.posterior.sample(n=n_samples, inla_model,
                                     num.threads = 5, selection=selection) %>%
    lapply(., function(x){
      x$latent
    }) %>% as.data.frame()

  colnames(inla_xg_posterior) <- paste0("sample", 1:n_samples)
  inla_xg_posterior$v <- rownames(inla_xg_posterior)
  rownames(inla_xg_posterior) <- c()
  inla_xg_posterior <- inner_join(inla_xg_posterior, rownames,
                                  by=c("v"="old_vars")) %>%
    select(-v)

  inla_xg_posterior_sample <- inla_xg_posterior %>% melt(id = "new_vars") %>%
    dplyr::select("new_vars", "value") %>%
    rename(vars = new_vars, sample = value)
}
xg_posterior_sample <- posterior(1000, inla_xg)
```

Plotting fixed effects posterior distributions

```{r}
xg_posterior_sample %>% filter(vars %in% c("x", "y", "shot_distance",
                                     "shot_angle", "seconds_since_last_shot")) %>%
  ggplot(aes(x=sample, group=vars)) + geom_density() +
  facet_wrap(~vars, scales="free")
```

Plotting posterior distributions for the top and bottom ten player random effects

```{r}
top_bottom_ten_scoring <- inla_xg$summary.random$shooter %>%
                       arrange(desc(mean)) %>% slice(c(1:10, (n()-10):n())) %>%
                        pull(ID)

xg_posterior_sample %>% filter(vars %in% top_bottom_ten_scoring) %>%
  ggplot(aes(x=sample, y=fct_reorder(vars, sample, .fun=mean))) +
  geom_density_ridges(quantile_lines = TRUE, quantiles = c(0.025, 0.5, 0.975))

```

```{r}
# posterior_predictive <- function(posterior_samples, variables){
#   posterior_samples %>% filter("(Intercept)") %>% as.vector()
# }
# fixed_variables <- list(x=0, y=0, shot_distance=1, shot_angle=1,
#     seconds_since_last_shot=1)
# random_variables <- list(player="AUSTON.MATTHEWS", goalie="TUUKKA.RASK",
#     event_detail="Snap", score_state="tied", game_strength_state="5v5")

fixed_effect_names <- c("x", "y", "shot_distance", "shot_angle",
                        "seconds_since_last_shot")
random_effect_names <- c("shooter", "goalie", "event_detail", "score_state",
                        "game_strength_state")





# post_pred <- xg_posterior_sample %>%
#   filter(vars == "(Intercept)") %>%
#   pull(sample)
#
# for(name in names(fixed_variables)){
#   post_pred <- post_pred + (xg_posterior_sample %>%
#                             filter(vars == name) %>%
#                             pull(sample)) * get(name, variables)
# }
#
# for(name in random_variables){
#   post_pred <- post_pred + (xg_posterior_sample %>%
#                             filter(vars == name) %>%
#                             pull(sample))
# }

gid <- "2020020001"
post_pred <- 0
for(name in fixed_effect_names){
  print(xg_data %>% filter(game_id == gid) %>% pull(name))
 post_pred <- post_pred + xg_data %>% filter(game_id == gid) %>% pull(name) %*%
    t(xg_posterior_sample %>% filter(vars == name) %>% pull(sample))
}
for(name in random_effect_names){
  post_pred <- post_pred + xg_data %>% filter(game_id == gid) %>% pull(name) %>%
  lapply(., function(x){
    # print(x)
    xg_posterior_sample %>% filter(vars == x) %>% pull(sample)
  }) %>% unlist(.) %>% matrix(., byrow=T, nrow=length(.)/1000)
}



# x <- inla.posterior.sample(n=1000, inla_xg,
#                                      num.threads = 5, selection=list("Predictor"=1:10)) %>%
#     lapply(., function(x){
#       x$latent
#     }) %>% as.data.frame()
#
# apply(x, 1, plogis) %>% apply(., 2, mean)
# apply(plogis(post_pred), 1, mean)
```

```{r}
fix_effect_samples <- list(
  intercept = xg_posterior_sample %>% filter(vars == "(Intercept)") %>% pull(sample),
  x = xg_posterior_sample %>% filter(vars == "x") %>% pull(sample),
  y = xg_posterior_sample %>% filter(vars == "y") %>% pull(sample),
  shot_angle = xg_posterior_sample %>%
    filter(vars == "shot_angle") %>% pull(sample),
  shot_distance = xg_posterior_sample %>%
    filter(vars == "shot_distance") %>% pull(sample),
  seconds_since_last_shot = xg_posterior_sample %>%
    filter(vars == "seconds_since_last_shot") %>% pull(sample)
)

gid <- "2020020001"
sub_xg_data <- xg_data %>% filter(game_id == gid)
post_pred <- apply(sub_xg_data, 1, function(row){
    fix_effect_samples$intercept +
    as.numeric(row[["x"]])*fix_effect_samples$x +
    as.numeric(row[["y"]])*fix_effect_samples$y +
    as.numeric(row[["shot_angle"]])*fix_effect_samples$shot_angle +
    as.numeric(row[["shot_distance"]])*fix_effect_samples$shot_distance +
    # as.numeric(row[["seconds_since_last_shot"]])*fix_effect_samples$shot_distance +
    xg_posterior_sample %>% filter(vars == row[["goalie"]]) %>% pull(sample) +
    xg_posterior_sample %>% filter(vars == row[["shooter"]]) %>% pull(sample) +
    xg_posterior_sample %>% filter(vars == row[["event_detail"]]) %>% pull(sample) +
    xg_posterior_sample %>% filter(vars == row[["score_state"]]) %>% pull(sample) +
    xg_posterior_sample %>% filter(vars == row[["game_strength_state"]]) %>% pull(sample)
})

apply(plogis(post_pred), 2, mean)
```


Fitting INLA model for each season
```{r}
for(s in pbp_data$season %>% unique()){
  sub_pbp <- pbp_data %>% filter(season == s)
  print(paste("Starting season:", s))
  xg_data <- data_pipeline_inla_xG(sub_pbp) %>% mutate(shooter = 
                      ifelse(Encoding(shooter) != "unknown", enc2native(shooter),
                          shooter))
  
  prior_random <- list(prec = list(prior = "normal", param = c(0, 1)))
  prior_fixed <- list(mean.intercept = 0, prec.intercept = 1,
                      mean = list(x = 0, y = 0, shot_distance = 0,
                                  shot_angle = 0, seconds_since_last_shot = 0),
                      prec = list(x = 1, y = 1, shot_distance = 1,
                                  shot_angle = 1, seconds_since_last_shot = 1))
  
  f <- isgoal ~ 1 + x + y + shot_distance + shot_angle + seconds_since_last_shot +
    f(goalie, model="iid", hyper = prior_random) +
    f(shooter, model="iid",  hyper = prior_random) +
    f(event_detail, model="iid",  hyper = prior_random) +
    f(score_state, model="iid", hyper = prior_random) +
    f(game_strength_state, model="iid", hyper = prior_random)
  
  print("Fitting Model")
  inla_xg <- inla(f, data=xg_data,
                      family="binomial",
                  control.predictor = list(link = 1),
                  control.fixed = prior_fixed,
                  control.compute = list(config=TRUE))
  print("Saving Model")
  saveRDS(inla_xg, paste0("models/xg_goal_hier_model_", s, ".RDS"))
  print("Finished")
}
```





