---
title: "Win Probability Model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create a dataframe with each row being a single nhl game, the outcome variable is an indicator variable that is 1 if the home team wins the game and 0 if the away team wins. The variables are

5v5 shots for rate, home team (over the last n games)
5v5 shots against rate, home team (over the last n games)
5v5 shots for rate, away team (over the last n games)
5v5 shots against rate, away team (over the last n games)

5v5 save percentage, home team (over the last n games)
5v5 save percentage, away team (over the last n games)

5v4 shots for rate, home team (over last n games)
5v4 shots for rate, away team (over last n games)

4v5 shots against rate, home team (over last n games)
4v5 shots against rate, away team (over last n games)

5v5 shooting %, home team (over last n games)
5v5 shooting %, away team (over last n games)

Indicator for home ice

days rest, home team
days rest, away team

Somehow include player talent into the model

Use a baseline model, which is home team always wins

```{r}
library(tidyverse)
events_summary <- read_csv("../data/events_summary_df.csv")
game_info <- read_csv("../data/game_info_df.csv")
pbp_base <- read_csv("../data/pbp_base.csv")
pbp_extras <- read_csv("../data/pbp_extras.csv")
player_periods <- read_csv("../data/player_periods.csv")
player_shifts <- read_csv("../data/player_shifts.csv")
report <- read_csv("../data/report.csv")
roster <- read_csv("../data/roster_df.csv")
scratches <- read_csv("../data/scratches_df.csv")
```

```{r}
# Gets 5v5, 5v4, 4v5 shots for per 60 for and shooting % for the home and away team
# Takes a dataframe which is the play-by-play data for a single game, scraped by Evolving Hockey's data scraper
shot_data_pipeline <- function(pbp){
  pbp %>% mutate(toi = max(game_seconds)) %>%
          filter(event_type == "SHOT" |
                 event_type == "GOAL") %>%
          filter(game_strength_state == "5v5" |
                 game_strength_state == "5v4" |
                 game_strength_state == "4v5") %>%
          mutate(x = abs(coords_x),
                 y = coords_y,
                 is_even = 1*(game_strength_state == "5v5"),
                 is_pp = 1*(game_strength_state == "5v4"),
                 is_sh = 1*(game_strength_state == "4v5"),
                 is_goal = 1*(event_type == "GOAL"),
                 is_even_goal = 1*(event_type == "GOAL" & game_strength_state == "5v5"),
                 team_type = ifelse(event_team == away_team, "AWAY", "HOME")) %>%
          group_by(team_type, event_team) %>%
          summarise(even_shots = sum(is_even),
                    pp_shots = sum(is_pp),
                    sh_shots = sum(is_sh),
                    time = max(toi),
                    ev_goals = sum(is_even_goal),
                    goals = sum(is_goal),
                    even_shots_per_60 = even_shots / (time / 60),
                    pp_shots_per_60 = pp_shots / (time / 60),
                    sh_shots_per_60 = sh_shots / (time / 60),
                    ev_shooting_percentage = ev_goals / even_shots,
                    save_percentage = (even_shots + pp_shots + sh_shots)/
                      (even_shots + pp_shots + sh_shots + goals)) %>%
          dplyr::select(team_type,
                        event_team,
                        even_shots_per_60,
                        pp_shots_per_60,
                        sh_shots_per_60,
                        save_percentage) %>%
          as_tibble(.)
}

x <- pbp_base %>%
      group_by(game_id) %>%
      nest() %>% 
      mutate(shot_data=map(.f=shot_data_pipeline(.))) %>% 
      unnest()

```